Handling String-number Conversion in String Constraint Solving

### 2019-09-26 draft.tex
some update

Such product automaton uses $Q\times P$ as the set of states and $\cvar^1_\epsilon \times \cvar^2_\epsilon$ as the alphabet. If a transition $((q_1,p_1), (x_{i,j},x'_{i',j'}),(q_2,p_2))$ is taken, for some $i,j,i',j' \in \mathbb{N}$, it means the character variable $x_{i,j}$ and $y_{i',j'}$ should be assigned the same value and this value leads $S_n$ from state $q_1$ to state $q_2$ and $S_m$ from $p_1$ to $p_2$.

@@

\begin{itemize}
\item If $(q_1,x_{i,j},q_2) \in T$, then for all $p\in P,((q_1,p),(x_{i,j},\epsilon),(q_2,p))\in T^2$.
\item If $(p_1,x_{i,j},p_2) \in T'$, then for all $q\in Q,((q,p_1),(\epsilon,x_{i,j}),(q,p_2))\in T^2$.
\item If $(q_1,x_{i,j},q_2) \in T$ and $(p_1,x'_{i',j'},p_2) \in T'$, then $((q_1,p_1),(x_{i,j},x'_{i',j'}),(q_2,p_2))\in T^2$.
\end{itemize}	
The first two cases correspond to the situation when the character variable $x_{i,j}$ is assigned $\epsilon$. In such cases, only one automaton changes its state.

For example, in the product automaton $A_{xy=yx}$ corresponds to $xy=yx$, we have the following transitions
\begin{itemize}
	\item $((q_0,p_0), (x_{0,0},\epsilon),(q_1,p_0))$, because $(q_0,x_{0,0},q_1)$ is a transition of $A_{xy}$.
	\item $((q_0,p_0), (\epsilon,y_{1,0}),(q_0,p_1))$, because $(p_0,y_{1,0},p_1)$ is a transition of $A_{yx}$.
	\item $((q_0,p_0), (x_{0,0},y_{1,0}),(q_1,q_1))$, because $(q_0,x_{0,0},q_1)$ is a transition of $A_{xy}$ and $(p_0,y_{1,0},p_1)$ is a transition of $A_{yx}$.
\end{itemize}

Recall that for flat automata, the number of occurrences of each symbol uniquely characterizes an accepting string (Section\ref{section:sfa}) and for any FA we can construct a Presburger formula characterizing its Parikh image (Section\ref{section:preliminary}). For a product automaton $A$ computed from the previous step, we compute a Presburger formula $\phi(A)$ over the set of variables $\{v_{(x,y)}\mid (x,y) \in \cvar_\epsilon \times \cvar_\epsilon\}$ characterizing the Parikh image of $A$. Then we add the following linear constraints.

\begin{enumerate}
	\item for any $(x_{i,j},y_{i',j'}) \in \cvar^1_\epsilon \times \cvar^2_\epsilon$, $v_{(x_{i,j},y_{i',j'})}>0 \rightarrow (x_j=y_{j})$. We use $-1$ to encode $\epsilon$.
	\item for all $x \in \cvar$, $|x| =\sum_{y \in \cvar_\epsilon} (v_{(x,y)}+v_{(y,x)})$
\end{enumerate}
Those constraints enforce that the two size of the equality of the word equation sharing the same interpretation.

We first build a product automata $\mathcal{P}=(Q\times P, T, \cvar\times \cvar, (q_i,p_i), (q_f,p_f))$, where $((q,p),(x,y), (q',p'))\in T$ iff $(q,x,q')\in T_Q \wedge (p,y,p')\in T_P$. Then the following formula characterizes all models of the equality constraint $Parikh(\mathcal{P}) \wedge \#(x,y) > 0 \rightarrow \enc{x} = \enc{y}$.

In practice, building the product automata is an expensive operation. In practice, we consider SFA of special shape that allows more efficient translation from equality constraint to linear integer constraint.

We define $(p,q)$-SFA as a SFA consisting of $p$ simple loops of length $q$, connected by epsilon transitions. Formally,

>>>>

Such product automaton uses $Q\times P$ as the set of states and $\cvar^1_\epsilon \times \cvar^2_\epsilon$ as the alphabet. If a transition $((q_1,p_1), (v_i^{(j)},v_{i'}^{(j')}),(q_2,p_2))$ is taken, for some $i,j,i',j' \in \mathbb{N}$, it means the character variable $v_i$ and $v_{i'}$ (and so all of their indexed variants) should be assigned the same value and this value leads $S_n$ from state $q_1$ to state $q_2$ and $S_m$ from $p_1$ to $p_2$.

@@

\begin{itemize}
\item If $(q_1,v_i^{(j)},q_2) \in T$, then for all $p\in P,((q_1,p),(v_i^{(j)},\epsilon),(q_2,p))\in T^2$.
\item If $(p_1,v_i^{(j)},p_2) \in T'$, then for all $q\in Q,((q,p_1),(\epsilon,v_i^{(j)}),(q,p_2))\in T^2$.
\item If $(q_1,v_i^{(j)},q_2) \in T$ and $(p_1,v_{i'}^{(j')},p_2) \in T'$, then $((q_1,p_1),(v_i^{(j)},v_{i'}^{(j')}),(q_2,p_2))\in T^2$.
\end{itemize}	
The first two cases correspond to the situation when the character variable $v_i^{(j)}$ is assigned $\epsilon$. In such cases, only one automaton changes its state.

For example, in the product automaton $A_{xy=yx}$ corresponds to $xy=yx$, we have the following transitions
\begin{itemize}
	\item $((q_0,p_0), (v_0^{(0)},\epsilon),(q_1,p_0))$, because $(q_0,v_0^{(0)},q_1)$ is a transition of $A_{xy}$.
	\item $((q_0,p_0), (\epsilon,v_2^{(1)}),(q_0,p_1))$, because $(p_0,v_2^{(1)},p_1)$ is a transition of $A_{yx}$.
	\item $((q_0,p_0), (v_0^{(0)},v_2^{(1)}),(q_1,q_1))$, because $(q_0,v_0^{(0)},q_1)$ is a transition of $A_{xy}$ and $(p_0,v_2^{(1)},p_1)$ is a transition of $A_{yx}$.
\end{itemize}

Recall that for flat automata, the number of occurrences of each symbol uniquely characterizes an accepting string (Section\ref{section:sfa}) and for any FA we can construct a Presburger formula characterizing its Parikh image (Section\ref{section:preliminary}). For a product automaton $A$ computed from the previous step, we compute a Presburger formula $\phi(A)$ over the set of variables $\{P{(v_i^{(j)},v_{i'}^{(j')})}\mid (v_i^{(j)},v_{i'}^{(j')}) \in \cvar^1_\epsilon \times \cvar^2_\epsilon\}$ characterizing the Parikh image of $A$. We use $P(v_i)$ and $P(v_i^{(j)})$ to denote the number of occurrences of $v_i$ and $v_i^{(j)}$, respectively. Then if $v_i^{(j)} \in \cvar^1$, then 
\begin{enumerate}
	\item $P(v_i^{(j)}) = \sum_{v \in \cvar^2_\epsilon} P{(v_i^{j},v)}$, if $v_i^{(j)} \in \cvar^1$  
	\item $P(v_i^{(j)}) = \sum_{v \in \cvar^1_\epsilon} P{(v,v_i^{j})}$, if $v_i^{(j)} \in \cvar^2$ 
\end{enumerate}
Note that $\cvar^1 \cap \cvar^2 =\emptyset$, because all variables will be renamed to some fresh variable. 
Then we add the following linear constraints.

\begin{enumerate}
	\item For any $(v_i^{(j)},v_{i'}^{(j')}) \in \cvar^1 \times \cvar^2$, 
	\begin{enumerate}
		\item $P{(v_i^{(j)},v_{i'}^{(j')})}>0 \rightarrow (v_i=v_{i'})$,
		\item $P{(v_i^{(j)},\epsilon)}>0 \rightarrow (v_i=\epsilon)$, and
		\item $P{(\epsilon,v_{i'}^{(j')})}>0 \rightarrow (v_{i'}=\epsilon)$.
	\end{enumerate}
	\item $P(v_i) = P(v_i^{(j)})$, for all $v_i\in \cvar$ and all different version numbers $j$ for $v_i$.
\end{enumerate}
The first constraint enforces the condition 1. and 2. and the second constraint enforces the condition 3. required for the pair of symbolic string $(w_n, w_m)$ and $I$. From the model $M$ of the linear constraints, we can obtain $w_n$, $w_m$, and $I$. $I(v)=M(v)$ for all $v\in \cvar$. Assume that the SFA $S_n$ is the one in Figure~\ref{fig:sfa_def}, then $$w_n = (v_0^0v_0^1v_0^2v_0^3)^{P(v_0^0)}v_0(v_1^0v_1^1v_1^2)^{P(v_1^0)}v_1\ldots(v_m^0v_m^1v_m^2v_m^3)^{P(v_m^0)}$$
One can construct $w_m$ in a similar way.
